---
title: "Asylum Applications in Europe"
author: "Sam Edwardes"
date: "2019-07-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

This chart was inspired by the [Makeover Monday 2019/W28 challenge](https://data.world/makeovermonday/2019w28). The data can be found at the same link. The original data source is from [Eurostat](http://appsso.eurostat.ec.europa.eu/nui/show.do?dataset=migr_asyappctza&lang=en), and the original visualization is from the [European Asylum Support Office (EASO)](https://www.easo.europa.eu/asylum-trends-annual-report-2018)

## Getting and Cleaning the data

The following code describes the necessary steps to get and clean the data before plotting.

```{r, environment, results=FALSE, warning=FALSE, message=FALSE}
setwd("~/Documents/GitHub/courses/09_DevelopingDataProducts/week 3 - assignment")
library("tidyverse")
library("httr")
library("readxl")
library("plotly")
```

```{r load_data, results=FALSE, warning=FALSE, message=FALSE}
file_name <- "eu_asylum.xls"
if(!file.exists(file_name)){
  download.file(url = "https://query.data.world/s/bgxvtr2aubwuin35e2tkkno2chbew5", destfile = file_name)
}
df_raw <- read_excel(file_name)
```

```{r clean_data, results=FALSE, warning=FALSE, message=FALSE}
df <- df_raw

# promote the headers
colnames(df) <- as.character(unlist(df[1,]))
df <- df[-1,]

# remove total lines
df <- df[c(-1, -34, -35, -68, -69, -102, -103, -136, -137, -170),]

# remove the columns with no information
df <- df %>% select(-SEX, -AGE)

# remove the rows where DECISION == "Total"
df <- df[df$DECISION != "Total",]

# transform from "long" to "tall" data
df <- df %>% gather(key = "year", value = "applications", `2009`:`2018`)

# fix data types
df$year <- as.numeric(df$year)
df$applications <- as.numeric(df$applications)

# Change the name of the GEO/TIME column
names(df)[1] <- "country"

# shorten "Germany (until 1990 former territory of the FRG)" to just "Germany"
df$country <- case_when(df$country == "Germany (until 1990 former territory of the FRG)" ~ "Germany",
                        TRUE ~ df$country)
```

## Assylum Applications Rejected by Country

```{r rejections_plot, warning=FALSE}
reject_country <- df[df$DECISION == "Rejected",]

plot_ly(data = reject_country, x = ~year, y = ~applications, 
        type = "scatter", mode = "lines",
        color = ~country) %>%
  layout(title = "Assylum Applications Rejected by Country",
         yaxis = list(title = "rejected applications"))
```

## Assylum Applications Accepted by Country

```{r accepted_plot, warning=FALSE}
accpet_country <- df[df$DECISION == "Total positive decisions",]

plot_ly(data = accpet_country, x = ~year, y = ~applications, 
        type = "scatter", mode = "lines",
        color = ~country) %>%
  layout(title = "Assylum Applications Accepted by Country",
         yaxis = list(title = "accepted applications"))
```

## Assylum Applications Acceptance Rate by Country

```{r accepted_rate_plot, warning=FALSE}
# keep only the accepted and rejected records
accept_rate <- df[df$DECISION == "Total positive decisions" | df$DECISION == "Rejected",]
accept_rate <- accept_rate %>% spread(key = DECISION, value = applications)
accept_rate <- accept_rate %>%
  mutate(total = `Rejected` + `Total positive decisions`,
         `Accepted Rate` = `Total positive decisions` / `total`)

plot_ly(data = accept_rate, x = ~year, y = ~`Accepted Rate`, 
        type = "scatter", mode = "lines",
        color = ~country) %>%
  layout(title = "Assylum Applications Acceptance Rate by Country",
         yaxis = list(title = "Rate of applicated accepted"))
```



